---
title: "Assemble and Upload pre-processed data"
author: "Rainer M. Krug"
format: 
  html:
    toc: true
    toc-location: left
    code-overflow: scroll
    code-fold: true
    code-tools: true
params:
  stage: [pre_processed]
  cores: 7
  archive_dir: "/Volumes/LEEF/LEEF-1.archived.data/"
  data_package_dir: "/Volumes/LEEF/LEEF-1 data packages"
  assemble: FALSE
  publish: FALSE
  sandbox: TRUE
---


## Introduction

This document assembles and publishes the pre-processed data and adds the metadata as defined in the function `metadata()` defined below.

Depending on the settings of `assemble` and `publish`, the data is assembled and / or published in the Zenodo Sandbox (`sandbox: TRUE`) or in the Zenodo repository. This should only be enabled if absolutely sure that everything is complete for publication. Data will be published as **Embargoed** until **WHICH TIME?**.

```{r}
#| label: setup

library(zenDep)
```

The following parameter will be used (explained later in the document):

```{r}
params
```
## Define metadata

The function `metadata()` must not require any parameter and returns a list containing the metadata.
```{r}
#| label: define_metadata

metadata <- new_metadata_bib(
  community = "leef-uzh",
  upload_type = "dataset",
  authors = list(
    list( firstname = "Romana",    lastname = "Limberger",         affiliation = "University of Zurich", orcid = "0000-0002-9421-7520"),
    list( firstname = "Uriah",     lastname = "Daugaard",          affiliation = "University of Zurich", orcid = "0000-0003-4092-717X"),
    list( firstname = "Rainer M.", lastname = "Krug",              affiliation = "University of Zurich", orcid = "0000-0002-7490-0066"),
    list( firstname = "Owen L.",   lastname = "Petchey",           affiliation = "University of Zurich", orcid = "0000-0002-7724-1633")
  ),
  description = "Description of the data",
  version = "1.0.0",
  language = "eng",
  keywords = c("LEEZ-UZH", "LEEF-1"),
  access_right = "embargoed",
  embargo_date = as.Date("2030-01-13"),
  license = "CC-BY-SA-4.0",
  contributors = list(
    list( firstname = "Anubhav",   lastname = "Gupta",             affiliation = "University of Zurich", orcid = "0000-0001-7977-9747", type = "DataCollector"),
    list( firstname = "Debra",     lastname = "Zuppinger-Dingley", affiliation = "University of Zurich", orcid = "0000-0002-4573-0563", type = "DataCollector"),
    list( firstname = "Frank",     lastname = "Pennekamp",         affiliation = "University of Zurich", orcid = "",                    type = "DataCollector"),
    list( firstname = "Martina",   lastname = "Jelic",             affiliation = "University of Zurich", orcid = "",                    type = "DataCollector"),
    list( firstname = "Owen L.",   lastname = "Petchey",           affiliation = "University of Zurich", orcid = "0000-0002-7724-1633", type = "ProjectLeader"), # << types TODO
    list( firstname = "Rainer M.", lastname = "Krug",              affiliation = "University of Zurich", orcid = "0000-0002-7490-0066", type = c("DataCurator", "DataManager")),
    list( firstname = "Romana",    lastname = "Limberger",         affiliation = "University of Zurich", orcid = "0000-0002-9421-7520", type = "ProjectLeader"), # << types TODO
    list( firstname = "Sabina",    lastname = "Jyrkinen",          affiliation = "University of Zurich", orcid = "",                    type = "DataCollector"),
    list( firstname = "Seraina",   lastname = "Nohl",              affiliation = "University of Zurich", orcid = "",                    type = "DataCollector"),
    list( firstname = "Sofia J.",  lastname = "van Moorsel",       affiliation = "University of Zurich", orcid = "0000-0003-1947-8971", type = "DataCollector"),
    list( firstname = "Uriah",     lastname = "Daugaard",          affiliation = "University of Zurich", orcid = "0000-0003-4092-717X", type = "ProjectLeader"), # << types TODO
    list( firstname = "Yves",      lastname = "Choffat",           affiliation = "University of Zurich", orcid = "",                    type = "DataCollector"),
    list( firstname = "Zheng",     lastname = "Xue",               affiliation = "University of Zurich", orcid = "",                    type = "DataCollector")
  ),
  grants = "10.13039/501100001711::310030_188431"
)

```

which will result in the following metadata for the deposits:

```{r}
metadata
```

## Assemble data packages
Now the data packages will be assembled and named accordingly if `parms$assemble == TRUE`. The original data is located in ``r params$archive_dir`` and saved in the directory ``r params$data_package_dir``. 
All timestamps in ``r params$archive_dir`` will be processed using ``r params$cores`` cores as specified in the parameter `params$cores`.

If `parms$assemble == FALSE`, nothing will be done and the data packaged will be expected in the directory ``r params$data_package_dir``.
```{r}
#| label: assemble_data_packages

if (params$assemble) {
  leef_create_data_archives(
    to_dir = params$data_package_dir,
    archive_dir = params$archive_dir,
    stage = params$stage,
    cores = params$cores
  )
}  
```

## Assemble data deposits and publish to Zenodo
This syep finally combines the data deposits for the pre-processed data (metadata & data package) and the extracted data (metadata & data package), uploads them to Zenodo (if `params$sandbox == TRUE` to the sandbox, the value is ``r params$sandbox``), and adds the corresponding related DOIs.

### Timestamps
The following timestamps will be deposited
```{r}
#| label: determine_timestamps

timestamps <- list.files(params$data_package_dir) |>
  (\(.) substr(., start = nchar(.)-11, stop = nchar(.) - 4))() |>
  unique()
timestamps
```

### Check pre-processed data packages
```{r}
#| label: check_pre_processed 
OK <- TRUE

fe <- file.path(params$data_package_dir, paste0("data_pre_processed_", timestamps, ".zip")) |>
  file.exists()
if (all(fe)) {
  pp_missing <- character(0)
} else {
  pp_missing <- paste0("data_pre_processed_", timestamps[!fe], ".zip")
  print("Missing pre_processed data packages!")
  print(pp_missing)
  OK <- FALSE
}
rm(fe)
```

### Check extracted data packages
```{r}
#| label: check_extracted

fe <- file.path(params$data_package_dir, paste0("data_extracted_", timestamps, ".zip")) |>
  file.exists()
if (all(fe)) {
  ex_missing <- character(0)
} else {
  ex_missing <- paste0("data_extracted_", timestamps[!fe], ".zip")
  print("Missing extracted data packages!")
  print(ex_missing)
  OK <- FALSE
}
rm(fe)
```

### Do the deposit
```{r}
#| label: deposit 
if (OK) {
  pbmcapply::pbmclapply(
    X = timestamps,
    FUN = zen_deposit,
    token = NULL,
    data_dir = params$data_package_dir,
    metadata_bib = metadata(),
    publish = params$publish,
    sandbox = params$sandbox,
    mc.cores = params$cores
  )
}
```

